---
title: "index"
author: "Anne Goj"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(sparkline)
library(reactable)
library(table1)
library(lubridate)
library(reactablefmtr)
library(htmltools)
library(flextable)
```
## Homework 4 Tables

## Data Source
https://www.gsa.gov/travel/plan-book/per-diem-rates/per-diem-files

## Summary
The U.S. General Services Administration sets rates for travel reimbursement annually for both lodging and for meals and incidental expenses. In 2023, the standard reimbursement for lodging will be \$98 and the standard reimbursement for meals will be \$59.  Typically a large metropolis or a popular tourist destination will have its own per Diem that may vary depending on the season.

I have taken the per diem files created tables from multiple years to show what the rates would be during a holiday in the Spring (St. Patrick's Day), the Summer (Independence Day), and the Fall (Thanksgiving Day). I create a reactable table that is filterable by year to show what the most expensive destinations may be.

## Data Manipulation

I upload the data and reformat the dollar columns as numbers. I calculate the difference between the destination's lodging rate and the standard US lodging rate.  For destinations without seasonal rate dependence, I set the season to start on January 1st and end on December 31. I use the lubridate package to convert the dates for seasonal rates into time intervals that the holidays may fall into.  

To compare data from different years, I move these data manipulations into a function.  I run the function on data from 2020-2022 and combine the rows into a large tibble.
```{r}
# get the standard rates for continental US over all years
standardrates <- read.csv("conusSTANDARDrates.csv",header=TRUE)

perdiem2023 <- read.csv("FY2023_PerDiemMasterRatesFile.csv",header=TRUE,skip=2)

newnames <- c("index","State","Destination","County","Season.Start","Season.End","Lodging","MandIE") 
names(perdiem2023) <- newnames

perdiem2023 %<>% mutate(Lodging=as.integer(str_remove(Lodging,"[ $]"))) %>% 
  mutate(MandIE=as.integer(str_remove(MandIE,"[ $]"))) %>% select(-index) %>% 
  mutate(State=as.factor(State)) 
USAdefaultLodging <- (standardrates %>% filter(year==2023))$standardLodging
USAdefaultMandIE <- (standardrates %>% filter(year==2023))$standardMIE
perdiem2023 %<>% mutate(LdiffUSDefault=Lodging-USAdefaultLodging)
perdiem2023 %<>% mutate(MdiffUSDefault=MandIE-USAdefaultMandIE)
perdiem2023 %>% head(2)
```

```{r}
perdiem2023%<>% mutate(Season.Start=if_else(Season.Start=="","January 1",Season.Start))
perdiem2023%<>% mutate(Season.End=if_else(Season.End=="","December 31",Season.End))

perdiem2023%<>% mutate(Season.End=mdy(paste(Season.End," 2023")))
perdiem2023%<>% mutate(Season.Start=mdy(paste(Season.Start," 2023")))
perdiem2023 %<>% mutate(Season=interval(Season.Start,Season.End))

baddates1 <- perdiem2023 %>% filter(Season<0)
baddates2 <- perdiem2023 %>% filter(Season<0)
perdiem2023 %<>% filter(Season>0)

baddates1 %<>% mutate(Season.End=mdy("December 31 2023")) %>%  mutate(Season=interval(Season.Start,Season.End))
baddates2 %<>% mutate(Season.Start=mdy("January 1 2023")) %>%  mutate(Season=interval(Season.Start,Season.End))

PD2023 <- bind_rows(perdiem2023,baddates1,baddates2) %>% arrange(State,Destination,County,Season.Start)

PD2023 %<>% mutate(Holiday="NOT")
dfID <- PD2023 %>%   mutate(Holiday=if_else(ymd("2023-07-4")%within%Season,"Independence Day",Holiday)) %>% dplyr::filter(Holiday!="NOT")
dfTD <- PD2023 %>%   mutate(Holiday=if_else(ymd("2023-11-24")%within%Season,"Thanksgiving Day",Holiday)) %>% dplyr::filter(Holiday!="NOT")
dfPD <- PD2023 %>%   mutate(Holiday=if_else(ymd("2023-3-17")%within%Season,"St. Patricks Day",Holiday)) %>% dplyr::filter(Holiday!="NOT")
PD2023 <- bind_rows(dfID,dfTD,dfPD) %>% arrange(State,Destination,County,Season.Start)
PD2023 %<>% mutate(Holiday=as.factor(Holiday))
PD2023 %<>%mutate(year=2023) 
```


```{r}

convert_PD_df <- function(mydf,myyear){
 newnames <- c("index","State","Destination","County","Season.Start","Season.End","Lodging","MandIE") 
names(mydf) <- newnames

mydf %<>% mutate(Lodging=as.integer(str_remove(Lodging,"[ $]"))) %>% 
  mutate(MandIE=as.integer(str_remove(MandIE,"[ $]"))) %>% select(-index) %>% 
  mutate(State=as.factor(State)) 
USAdefaultLodging <- (standardrates %>% filter(year==myyear))$standardLodging
USAdefaultMandIE <- (standardrates %>% filter(year==myyear))$standardMIE
mydf %<>% mutate(LdiffUSDefault=Lodging-USAdefaultLodging)
mydf %<>% mutate(MdiffUSDefault=MandIE-USAdefaultMandIE)
mydf %>% head(2)
mydf%<>% mutate(Season.Start=if_else(Season.Start=="","January 1",Season.Start))
mydf%<>% mutate(Season.End=if_else(Season.End=="","December 31",Season.End))

mydf%<>% mutate(Season.End=mdy(paste(Season.End,myyear)))
mydf%<>% mutate(Season.Start=mdy(paste(Season.Start,myyear)))
mydf %<>% mutate(Season=interval(Season.Start,Season.End))

baddates1 <- mydf %>% filter(Season<0)
baddates2 <- mydf %>% filter(Season<0)
mydf %<>% filter(Season>0)

yearend=paste("December 31",myyear)
yearstart=paste("January 1",myyear)
baddates1 %<>% mutate(Season.End=mdy(yearend)) %>%  
  mutate(Season=interval(Season.Start,Season.End))
baddates2 %<>% mutate(Season.Start=mdy(yearstart)) %>%
  mutate(Season=interval(Season.Start,Season.End))
mynewPD <- bind_rows(mydf,baddates1,baddates2) %>%
  arrange(State,Destination,County,Season.Start)

mynewPD %<>% mutate(Holiday="NOT")
july4=paste("July 4",myyear)
thankgiv <- paste("November 24",myyear)
stpaddy <- paste("March 17",myyear)
dfID <- mynewPD %>%   mutate(Holiday=if_else(mdy(july4)%within%Season,
                                             "Independence Day",Holiday)) %>%
  dplyr::filter(Holiday!="NOT")
dfTD <- mynewPD %>%   mutate(Holiday=if_else(mdy(thankgiv)%within%Season,
                                             "Thanksgiving Day",Holiday)) %>%
  dplyr::filter(Holiday!="NOT")
dfPD <- mynewPD %>%   mutate(Holiday=if_else(mdy(stpaddy)%within%Season,
                                             "St. Patricks Day",Holiday)) %>%
  dplyr::filter(Holiday!="NOT")
mynewPD <- bind_rows(dfID,dfTD,dfPD) %>% 
  arrange(State,Destination,County,Season.Start)
mynewPD %<>% mutate(Holiday=as.factor(Holiday))
mynewPD %<>%mutate(year=as.integer(myyear)) 
return(mynewPD)
}

perdiem2022 <- read.csv("FY2022_PerDiemMasterRatesFile.csv",header=TRUE,skip=2)
perdiem2021 <- read.csv("FY2021_PerDiemMasterRatesFile.csv",header=TRUE,skip=2)
perdiem2020 <- read.csv("FY2020_PerDiemMasterRatesFile.csv",header=TRUE,skip=2)

PD_allyears <- bind_rows(PD2023,convert_PD_df(perdiem2022,2022),
                         convert_PD_df(perdiem2021,2021),
                         convert_PD_df(perdiem2020,2020))
```


```{r}
PD_allyears %>% filter(Destination=="Fresno",Holiday=="Independence Day")
```


## Static Tables
I use table1 package to compare Lodging and Meals rates across different holidays for the 2023 data set.
Then I create a table comparing Summer and Fall rates between New York and Florida for the 2023 set.
```{r}
label(PD2023$MandIE) <- "Meals and Incidentals,USD"
label(PD2023$Lodging) <- "Lodging,USD"
tableA <- table1(~Lodging+MandIE|Holiday, data=PD2023 ,topclass="Rtable1-shade")
tableA
tableB <- table1(~Lodging+MandIE|Holiday*State, data=PD2023 %>% filter(State%in%c("FL","NY"),Holiday%in%c("Independence Day","Thanksgiving Day")))
tableB
```


## Reactable Table
I created a Reactable table to show the most expensive Destinations to travel. I aggregate to display the maximum lodging and the mean meals expenses for each for each state by year. I allow selection of the data by year. As you drill down into the state's destinations, there is a bar chart to show how much more expensive they would be compared to the standard lodging. 

```{r}

# PD_allyears%>% 
  # select(State, Destination, Lodging,LdiffUSDefault,MandIE, Holiday,year)
reactable_table <-PD_allyears %>%
  group_by(State,year) %>% 
  dplyr::select(State, Destination, Lodging,LdiffUSDefault,MandIE, Holiday,year) %>% 
  reactable(groupBy = c("year","State"),
            paginateSubRows = TRUE,
            filterable = TRUE,
            style = list(fontSize = "1rem"),columns = list(
    year=colDef(name="Year",filterable=TRUE,
             filterInput = function(values, name) {
        tags$select(
          # Set to undefined to clear the filter
          onchange = sprintf("Reactable.setFilter('year-select',
                             '%s', event.target.value || undefined)", name),
          # "All" has an empty value to clear the filter, and is the default option
          tags$option(value = "", "All"),
          lapply(unique(values), tags$option),
          "aria-label" = sprintf("Filter %s", name),
          style = "width: 100%; height: 28px;"
        ) },
                ),
    State=colDef(
      filterInput = function(values, name) {
        tags$select(
          # Set to undefined to clear the filter
          onchange = sprintf("Reactable.setFilter('year-select',
                             '%s', event.target.value || undefined)", name),
          # "All" has an empty value to clear the filter, and is the default option
          tags$option(value = "", "All"),
          lapply(unique(values), tags$option),
          "aria-label" = sprintf("Filter %s", name),
          style = "width: 100%; height: 28px;"
        ) }
    ),
   Destination=colDef(
      filterInput = function(values, name) {
        tags$select(
          # Set to undefined to clear the filter
          onchange = sprintf("Reactable.setFilter('year-select',
                             '%s', event.target.value || undefined)", name),
          # "All" has an empty value to clear the filter, and is the default option
          tags$option(value = "", "All"),
          lapply(unique(values), tags$option),
          "aria-label" = sprintf("Filter %s", name),
          style = "width: 100%; height: 28px;"
        ) }
    ),
    Lodging = colDef(aggregate = "max",filterable = FALSE,
                     format = list(cell=colFormat(currency = "USD"),
                                   aggregated = colFormat(prefix = "maximum =",
                                                          currency = "USD")),
                     name = "Lodging"),
    LdiffUSDefault = colDef(name = "Over Standard Lodging Rate",filterable = FALSE,
                            cell = data_bars(., fill_color = c("red","black"),
                                             fill_gradient = TRUE,
                                             number_fmt=scales::number_format(accuracy=1)
                                             )),
    MandIE = colDef(aggregate = "mean",filterable = FALSE,
                    format = list(cell=colFormat(currency = "USD"),
                                  aggregated = colFormat(prefix = "mean =",
                                                         currency = "USD")),
                    name = "Meals & Incidentals"),
    Holiday=colDef(
      filterInput = function(values, name) {
        tags$select(
          # Set to undefined to clear the filter
          onchange = sprintf("Reactable.setFilter('year-select',
                             '%s', event.target.value || undefined)", name),
          # "All" has an empty value to clear the filter, and is the default option
          tags$option(value = "", "All"),
          lapply(unique(values), tags$option),
          "aria-label" = sprintf("Filter %s", name),
          style = "width: 100%; height: 28px;"
        ) }
    )
    ),
    defaultSorted = "Lodging",
    defaultSortOrder = "desc",
    theme=clean(),
    defaultPageSize = 15,
    elementId = "year-select")  %>%   add_title("Per Diem Travel Rates ") %>% 
  add_subtitle("as issued by the US Goverment for various holidays.") %>% 
  add_source("https://www.gsa.gov/travel/plan-book/per-diem-rates/per-diem-files")
  
reactable_table
```

